### 2025-12-04 presentations/social_mobility_2025/mjp-mobility-conf-2025.qmd
Up ahora si
up

## Causal model (M): @breen_effects_2024

@@

## A Causal Model for Mobility Effects

@@

## Inquiry and estimand
::: {.incremental .highlight-last style="font-size: 120%;"}

@@

## Treatment: mobility and class matrix (D)

@@

## Weights and covariates (D → A)

@@

## Heterogeneity and controls (D → A)

@@

## Robustness check and sensitivity analysis

- **Unmeasured confounding** [@cinelli_making_2020]: **Moderate robustness** in Middle→High and High→Low; borderline in High→Middle.

- **Robustness check**: I re-estimate the models under three alternative codings of the outcome: (i) recoding the neutral category as agreement (3–5 vs. 1–2), (ii) excluding neutral responses from the analysis, and (iii) using the original 5-point item in a linear specification. The results are maintained and even improved (only in ii).
:::

@@

1) Most Chileans reject income-based pension allocation, yet support for this market principle has grown in a highly marketised and unequal setting [@lindh_public_2015; @castillo_perceptions_2025].

2) Intergenerational moves into and out of upper-status positions causally shape support for pension commodification, while donward High-Middle y High-Low decrease it [@gugushvili_subjective_2017; @rodriguez_intergenerational_2024; @ares_changing_2020]; meritocratic beliefs correlate with these preferences but seem to operate on a parallel track.

>>>>

## Mobility effects approach

@@

## Causal model (M)

@@

## Inquiry and estimand (I)
::: {.incremental .highlight-last style="font-size: 120%;"}

@@

## Treatment: mobility and class matrix

@@

## Weights and covariates

@@

## Heterogeneity and controls

@@

@@

1) Intergenerational moves into and out of upper-status positions causally shape support for pension commodification, while donward High-Middle y High-Low decrease it [@gugushvili_subjective_2017; @helgason_longterm_2023; @busemeyer_welfare_2020; @ares_changing_2020]

2) Meritocratic beliefs correlate with these preferences but seem to operate on a parallel track [@@rodriguez_intergenerational_2024]
